---
title: "salt_effect"
author: "Guy Mercer"
date: "27/11/2020"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

Machine Settings:

* Number of Cycles - 192

* Cycle Length - 300s

* Wavelength - 600nm

* Temperature - 30°C

* Shake - Before Each Cycle (258s)

* Shake Mode - Double Orbital

* Shake Frequency - 200rpm

* Total Well Volume - 200µl

Formatted microplate reader output file to remove gaps and make data easier to process. For example, edited time into minute format. The amount of shaking has changed to 258s because this was the maximum when 96 wells are being measured. The more wells measured, the longer it takes the spectrophotometer to take the measurements. 

```{zsh raw-file-formatting}
#!/usr/bin/env zsh

# setwd
cd /Users/guy/Documents/phd/DEG_yeast_insecticide_expt/senstivity_test/insecticides_salt_sens

./WHATIDID.sh 
```

```{r}
library(tidyverse)
library(knitr)
library(growthcurver)
library(pastecs)
library(ibd)
library(rstatix)
library(ggpubr)
library(broom)
library(patchwork)
library(sjPlot)
library(sjmisc)
library(sjlabelled)
# library(fitdistrplus)
``` 

```{r}
# set wd
setwd("~/Documents/phd/DEG_yeast_insecticide_expt/senstivity_test/insecticides_salt_sens/")

# read csv
insect_sens <- read.csv(file = "./results/insect_salt_compiled.csv", header = FALSE, stringsAsFactors = FALSE)

# formatting function
sens_data_form <- function(sens_data) {
  # transpose
  sens_data_trans <- t(sens_data)
  # save first row as colnames
  colnames(sens_data_trans) <- sens_data_trans [1, ]
  # delete first row, which is now duplicate of colnames
  sens_data_colnames <- sens_data_trans [-1, ]
  # remove any empty rows. Sometimes created when importing csv file 
  sens_data_no_blanks <- sens_data_colnames[rowSums(is.na(sens_data_colnames)) != ncol(sens_data_colnames),]
  # convert matrix into dataframe
  sens_data_df <- as.data.frame(sens_data_no_blanks)
  # convert everything to numeric
  sens_data_df[] <- lapply(sens_data_df, function(x) {
    if(is.factor(x)) as.numeric(as.character(x)) else x
  })
  # convert to data frame
  sens_data_corr <- as.data.frame(sens_data_df)
  
  # rename row names to numbers
  rownames(sens_data_corr) <- 1:length(rownames(sens_data_corr))
  
  return(sens_data_corr)
}

sens_data_corr <- sens_data_form(insect_sens)

# set any negative values to 0
sens_data_corr[sens_data_corr<0] <- 0
```

To input the data into Growthcurver I had to alter the formatting. When using Growthcurver, columns have to be looped through individually for a more detailed output. Growth curve data is fitted to the standard form of the logistic equation (sigmoidal). 

```{r, include = FALSE}
growthcurver_function <- function(sens_data_corr) {
  
# Let's create an output data frame to store the results in. 
# We'll create it so that it is the right size (it's faster this way!), 
# but leave it empty.
num_analyses <- length(names(sens_data_corr)) - 1
d_gc <- data.frame(sample = character(num_analyses),
                   k = numeric(num_analyses),
                   n0  = numeric(num_analyses),
                   r = numeric(num_analyses),
                   r_se = numeric(num_analyses),
                   r_p = numeric(num_analyses),
                   t_mid = numeric(num_analyses),
                   t_gen = numeric(num_analyses),
                   auc_l = numeric(num_analyses),
                   auc_e = numeric(num_analyses),
                   sigma = numeric(num_analyses),
                   stringsAsFactors = FALSE)

# Now, loop through all of the columns in the data frame. For each column,
# run Growthcurver, save the most useful metrics in the output data frame,
# and make a plot of all the growth curve data and their best fits.
par(mfcol = c(4,4))
par(mar = c(1,1,1,1))
y_lim_max <- max(sens_data_corr[,setdiff(names(sens_data_corr), "time")]) - min(sens_data_corr[,setdiff(names(sens_data_corr), "time")])

n <- 1    # keeps track of the current row in the output data frame
for (col_name in names(sens_data_corr)) {
  
  # Don't process the column called "time". 
  # It contains time and not absorbance data.
  if (col_name != "time") {
    
    # Create a temporary data frame that contains just the time and current col
    sens_data_corr_loop <- sens_data_corr[, c("time", col_name)]
    
    # Now, call Growthcurver to calculate the metrics using SummarizeGrowth
    gc_fit <- SummarizeGrowth(data_t = sens_data_corr_loop[, "time"], 
                              data_n = sens_data_corr_loop[, col_name],
                              bg_correct = "none")
    
    # Now, add the metrics from this column to the next row (n) in the 
    # output data frame, and increment the row counter (n)
    d_gc$sample[n] <- col_name
    d_gc[n, 2:11] <- c(gc_fit$vals$k,
                       gc_fit$vals$n0,
                       gc_fit$vals$r,
                       gc_fit$vals$r_se,
                       gc_fit$vals$r_p,
                       gc_fit$vals$t_mid,
                       gc_fit$vals$t_gen,
                       gc_fit$vals$auc_l,
                       gc_fit$vals$auc_e,
                       gc_fit$vals$sigma)
    n <- n + 1
    
    # Finally, plot the raw data and the fitted curve
    n_obs <- length(gc_fit$data$t)
    plot(gc_fit$data$t, gc_fit$data$N, 
         pch = 20, 
         xlim = c(0, 960), 
         ylim = c(0, y_lim_max),
         cex = 0.6, xaxt = "n", yaxt = "n")
    text(x = 960 / 2, y = y_lim_max, labels = col_name, pos = 1)
    lines(gc_fit$data$t, predict(gc_fit$model),col = "red")
  }
}

# Uncomment the next line to save the plots from your 96-well plate to a file
# dev.off()

return(d_gc)

}

d_gc <- growthcurver_function(sens_data_corr = sens_data_corr)
```

```{r}
replicates_of_interest <- colnames(sens_data_corr)

replicates_of_interest_df <- tibble(matrix(nrow=nrow(sens_data_corr)))

for (sample in 1:length(replicates_of_interest)) {
  
  replicate_col <- sens_data_corr [ , grep(pattern = replicates_of_interest [sample], x = names(sens_data_corr))]
  
  replicates_of_interest_df [ , sample] <- replicate_col
  
}

colnames(replicates_of_interest_df) <- replicates_of_interest 

replicates_of_interest_curves <- growthcurver_function(sens_data_corr = replicates_of_interest_df)

```

Unsure what caused the cloth-high-salt scatter at the beginning of some of the growth curves. 

```{r}
d_gc <- as_tibble(d_gc)

# Plot a histogram of the sigma values in order to check for outliers
hist(d_gc$sigma, main = "Histogram of sigma values", xlab = "sigma")

# Show the top 50 samples with the largest sigma value 
# (with the worst model fit to the growth curve data)
high_sigma <- d_gc %>% top_n(20, sigma) %>% arrange(desc(sigma))

high_sigma
```

```{r}
pca_gc_out <- as_tibble(d_gc) 

# Prepare the gc_out data for the PCA
rownames(pca_gc_out) <- pca_gc_out$sample

# Do the PCA
pca.res <- prcomp(pca_gc_out %>% select(k:sigma), center=TRUE, scale=TRUE)

# pdf(file="pca.pdf") if i want to save the plot
# Plot the results
as_tibble(list(PC1=pca.res$x[,1],
                   PC2=pca.res$x[,2],
                   samples = rownames(pca_gc_out))) %>% 
  ggplot(aes(x=PC1,y=PC2, label=samples)) + 
geom_text(size = 1)
# dev.off()
```

```{r}
# save a complete version of the data.
growthcurver_data <- d_gc

d_gc <-  d_gc [-grep(pattern = "control-salt-2-13", x = d_gc$sample), ]

d_gc <-  d_gc [-grep(pattern = "imi-high-2-8" , x = d_gc$sample), ]

```

Create a linear model with auc_l as the dependent variable and treatment as the factor with salt separately. 

```{r}
# extract sample and auc_l from d_gc
sample_aucl <- d_gc [, c(1,9)]

sample <- sample_aucl [, 1]

# add insecticide column 
factor_levels <- c("control", "acet", "cloth",
                   "flu", "imi")

for (level in 1:length(factor_levels)) {
  
  level_name <- paste0("^", factor_levels [level], "-.*$", sep = "")
  
  sample_aucl$sample <- gsub(pattern = level_name, replacement = factor_levels [level], x = sample_aucl$sample)

}

colnames(sample_aucl) <- c("insecticide", "auc_l")

# add sample column that conserves replicate number back to df
sample_aucl [, 3] <- sample

# add salt column 
salt_logical <- grepl(pattern = "*-salt-*", x = sample_aucl$sample)

sample_aucl$salt <- salt_logical

# add a column for day
sample_aucl [, 5] <- sample

colnames(sample_aucl) <- c("insecticide", "auc_l", "sample", "salt", "day")

day_names <- c("one", "two")

for (day in 1:length(day_names)) {
  
  day_name <- paste0(".*-", day, "-.*", sep = "")
  
  sample_aucl$day <- gsub(pattern = day_name, replacement = day_names [day], x = sample_aucl$day)

}

# add a conc column 
sample_aucl [, 6] <- sample

colnames(sample_aucl) <- c("insecticide", "auc_l", "sample", "salt", "day", "conc")

conc_names <- c("low", "high")

for (conc in 1:length(conc_names)) {
  
  conc_name <- paste0(".*-", conc_names [conc], "-.*", sep = "")
  
  sample_aucl$conc <- gsub(pattern = conc_name, replacement = conc_names [conc], x = sample_aucl$conc)

}

control_values <- sample_aucl$conc [grep(pattern = "control", x = sample_aucl$conc)]

for (i in 1:length(control_values)) {
  
  #sample_aucl <- sample_aucl %>% replace_with_na_at(.vars = "conc", condition = ~.x == control_values [i])
  
  sample_aucl$conc <- gsub(pattern = control_values [i], replacement = "low" , x = sample_aucl$conc )
}

# create a treatment vector with both treatments and treatments with salt
boxplot_treatments <- c("control", "acet-high", "acet-low", "cloth-high", "cloth-low",
                   "flu-high", "flu-low", "imi-high", "imi-low")


boxplot_treatments_salt <- boxplot_treatments

for (i in 1:length(boxplot_treatments)) {
  
  new_name <- paste0(boxplot_treatments [i], "-salt", sep = "")
  
  boxplot_treatments_salt <- gsub(pattern = boxplot_treatments [i], replacement = new_name, x = boxplot_treatments_salt)
  
}

boxplot_treatments_both <- c(boxplot_treatments, boxplot_treatments_salt)

# create column in sample_aucl with vector
for (treatment in 1:length(boxplot_treatments_both)) {
  
  treatment_name <- paste0("^", boxplot_treatments_both [treatment], "-\\d+-\\d+$", sep = "")
  
  sample_aucl$sample <- gsub(pattern = treatment_name, replacement = boxplot_treatments_both [treatment], x = sample_aucl$sample)

}

colnames(sample_aucl) <- c("insecticide", "auc_l", "boxplot_treatment", "salt", "day", "conc")


sample_aucl [, 7] <- sample

# add treatment column 
treatment_levels <- c("control", "acet-low", "acet-high", "cloth-low", "cloth-high",
                   "flu-low", "flu-high", "imi-low", "imi-high")

for (level in 1:length(treatment_levels)) {
  
  treatment_name <- paste0("^", treatment_levels [level], "-.*$", sep = "")
  
  sample_aucl$sample <- gsub(pattern = treatment_name, replacement = treatment_levels [level], x = sample_aucl$sample)

}

colnames(sample_aucl) <- c("insecticide", "auc_l", "boxplot_treatment", "salt", "day", "conc", "treatment")

sample_aucl [, 8] <- sample

# convert treatment to factor
sample_aucl$treatment <- as.factor(sample_aucl$treatment)

# reorder insecticide factor so control comes first
sample_aucl$treatment <- relevel(sample_aucl$treatment, "control")

# convert insecticide to factor
sample_aucl$insecticide <- as.factor(sample_aucl$insecticide)

# reorder insecticide factor so control comes first
sample_aucl$insecticide <- relevel(sample_aucl$insecticide, "control")

aucl_treatment_model <- lm(auc_l ~ (salt * treatment), data = sample_aucl)

summary(aucl_treatment_model)
```

```{r}
control <- sample_aucl [grep(pattern = "control", x = sample_aucl$insecticide), ]
control_salt <- control [grep(pattern = "^FALSE$", x = control$salt), ]
mean(control_salt$auc_l)
```


Day has no effect this time so has been left out of the model. 

Test assumptions of linear regression. In the model output above, the significance value for the intercept is testing whether the intercept is significantly different from zero. Rather useless in this situation. 

By removing control-salt-2-13 and imi-high-2-8 the model residuals pass a shapirowilks test. Removing them doesn't change the model much output either. 

```{r}
# Inspect the model diagnostic metrics
aucl_treatment_model_metrics <- augment(aucl_treatment_model)

# normality test
shapiro_result <- shapiro_test(aucl_treatment_model_metrics$.resid)

shapiro_result

# homogeneity of variance
par(mfrow = c(2, 2))
plot(aucl_treatment_model)

plot(density(aucl_treatment_model_metrics$.resid))
```

For aucl ~ treatment + salt + (salt * treatment) the output is essentially the same but now the synergy and discord is calculated for me. 

The interaction values show the difference between what would be expected if the effect of the two stressors in combination was additive and the actual result. For example, treatmentacet-high = 124.816 and salt = 5.833. Therefore, if additive treatmentacet-high = 130.649. Actual result was -157.349 lower than this, indicative of synergy.

Talking points. Why do acet and flu in isolation appear to stimulate growth? Is there any practical application for that? All insecticides at the high concentration (5mM) have a synergistic effect with salt. In the case of acet and flu, however, this isn't very useful because all this synergy does it reverse the seemingly stimulatory effect of the insecticide in isolation. 

Boxplot by Treatment

```{r}
# convert boxplot_treatment to factor
sample_aucl$boxplot_treatment <- as.factor(sample_aucl$boxplot_treatment)

# reorder insecticide factor so control comes first
sample_aucl$boxplot_treatment <- factor(sample_aucl$boxplot_treatment, levels = c("control", "control-salt", "acet-high", "acet-low", "acet-high-salt", 
                                                                                  "acet-low-salt","cloth-high", "cloth-low", "cloth-high-salt", "cloth-low-salt",
                                                                                  "flu-high", "flu-low", "flu-high-salt", "flu-low-salt", "imi-high", "imi-low", 
                                                                                  "imi-high-salt","imi-low-salt"))

boxplot(auc_l ~ boxplot_treatment, data = sample_aucl,
                las=2, par(mar = c(7, 4.1, 4.1, 2.1)),
                xlab = "")

```

Lots of spread in the control. 
